---
title: "Fun Programming Lab 3"
author: "Stephanie Gluck"
date: "5/9/2020"
output: 
  html_document: 
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
# Part A: Foundations

Briefly name and describe the three fundamental components of a function.  

## Three components of a function:  

1. body()  The statements that are executed by calling a particular function.

2. formals() The arguments supplied to the function.

3. environment()  The location of the function's variables - either a package or a global enviroment. 

Describe three different ways functions can be stored/applied and an example of when you might want to use each version.


1.  list - Can loop through several functions for a particular vector.

2.  object - bound to a name

3.  anonymous function - function not bound to a name

# Part B: Applied practice

The mean is defined as:
∑ni=1xin where xi represents each element in a generic vector, x, ∑ represents the summation of these elements, and n represents the total number of elements in x. In other words, sum and divide by the number included.

Write a function to calculate the mean that removes missing data before conducting the calculation. You may not use base::mean or any similar function that conducts a mean calculation.

```{r mean_function}

data <- c(1, 4, 6, NA, 8, NA, 10)

length(data)

#need to omit NA for length 
mean_fn <- function(x){
  y <- na.omit(x)
  sum(y)/length(y)
}

```

Test your function to make sure it (a) provides the expected results, and (b) gives identical output to base::mean when na.rm = TRUE. Make sure that you test your data against a vector that has missing data.
```{r test_function}

# a. testing the function we created
mean_fn(data)

# b. compare with base::mean function
mean(data, na.rm = TRUE)

```

Write a function that takes two vectors of the same length and returns the total number of instances where the value is NA for both vectors. For example, given the following two vectors (see code). Test that the function works as expected on two new vectors (i.e., any two vectors that are not given).
```{r vector_function}

a <- c(1, NA, NA, 3, 3, 9, NA)
b <- c(NA, 3, NA, 4, NA, NA, NA)

sum(is.na(a) &
is.na(b))

#function
na_match <- function(x, y) {
 sum(is.na(x) &
 is.na(y))
}

#test with a,b
na_match(a, b)

#test with new vectors that has 3 overlapping NA
c <- c(1,4,6,NA,8,NA,10,NA)
d <- c(2,4,6,NA,7,NA,9,NA) 

na_match(c, d)

```

Turn the following three lines of code into three different functions. Make sure to give them meaningful names. Test the functions to make sure they provide the expected output.
```{r count_missing}
# Count number of missing elements in a vector
sum(is.na(airquality$Ozone))

#function
count_missing <- function(x) {
  sum(is.na(x))
}

#test
count_missing(airquality$Ozone)

```

```{r prop_represent}
# Proportional representation of each level (unique element) in a vector
purrr::map_dbl(split(mtcars$cyl, mtcars$cyl), length) / length(mtcars$cyl)

#function
prop_represent <- function(x) {
  purrr::map_dbl(split(x, x), length) / length(x)
}

#test
prop_represent(mtcars$cyl)
prop_represent(mtcars$disp)

```

```{r normalize}
# normalize or z-score a vector so the mean is zero and sd is one
(mtcars$mpg - mean(mtcars$mpg, na.rm = TRUE)) / 
  sd(mtcars$mpg, na.rm = TRUE)

#function
normalize <- function(x) {
  (x - mean(x, na.rm = TRUE)) / 
  sd(x, na.rm = TRUE)
}

#test
normalize(mtcars$mpg)
```

Write a function that takes a data frame as its input and returns a data frame with only the numeric columns.
```{r df_numeric}

# returns a logical vector 
map_lgl(iris, is.numeric)

# use brackets  to subset 
iris[ , map_lgl(iris, is.numeric)]

# wrapping in function 
df_numeric <- function(df) {
  df[ , map_lgl(df, is.numeric)]
}

# checking 
df_numeric(iris)
df_numeric(diamonds)

```

Write a function that uses the function you wrote in Question 5, and returns a data frame with the mean and standard deviation of each numeric column.
```{r df_msd}

df_msd <- function(df) {
  df_num <- df[ , map_lgl(df, is.numeric)]
  means <- map(df_num, mean_fn)
  sd <- map(df_num, sd)
  return(as.data.frame(c(means = means, sd = sd)))
}

df_msd(iris)
df_msd(diamonds)

```

**Stephanie comment:** I wanted to try to find a way to "tidy up" the output that was generated by the function. However I could not figure out the regular experssion / `names_pattern` matching to remove the first period (.) after mean and sd

```{r}


df_msd <- function(df) {
  df_num <- df[ , map_lgl(df, is.numeric)]
  means <- map(df_num, mean_fn)
  sd <- map(df_num, sd)
  d <- as.data.frame(c(means = means, sd = sd))
  return(tidyr::pivot_longer(d, 
                             everything(),
                             names_to = c("descriptive", "variables"),
                             names_pattern = "([A-Za-z]+)(.*)",
                             values_to = "values") %>% 
           pivot_wider(names_from = "descriptive", 
                       values_from = "values"))
}

df_msd(iris)
df_msd(diamonds)

```

